### 2019-03-12 manuscript/discussion.tex
- backups of all the changes -- mainly to discussion

%B:  at least two surprising outcomes: (1)  D and P on background of C, and (2) SC not as dead of an end as it was in goldberg2010, if P is partitioned out (can be framed as kind of path-dependence, i.e. qIC path different than rho path, or same as (1), so that I and C are less different on the background of D.
We investigated the individual and joint effects of ploidy and breeding system states in diversification of Solanaceae.
Most significantly, we found that the effects of ploidy, inferred as significant when it is the only trait under consideration, are overshadowed by hidden traits, or breeding system when added to models.  
Interestingly, the difference in diversification rates between the diploids and polyploids is negligible on the background of self-compatibility.
In the same vein, we recovered a smaller difference in diversification rates between self-incompatible and self-compatible lineages against the diploid genomic background.
%Inference of transition rates can appreciably across 
Motivated by the widespread findings of paleopolyploid ancestry across angiosperms, we also examined and found modest support for diploidization in the history of the family. %B: "mixed" "not strong" what's the right word here?  E: not "weak", maybe "not strong" or "not nearly as strong as anything else"
Disentangling the patterns of diversification linked to multiple traits is difficult, but our strategy allows tests of intricate hypotheses regarding a heterogenous diversification process.
Below, we discuss these key findings and detail their sometimes confusing and complex implications.
We additionally provide an outline for broader application of phylogenetic comparative methods in the context of interactions among traits and diversification process.

% - value of approach
Disentangling the patterns of diversification linked to  multiple traits allowed us the opportunity to model and test complex hypotheses about the heterogenous diversification process in Solanaceae.
Species are created and go extinct based on multiple and often highly correlated phenotypes. However, estimating the speciation and extinction rates remains challenging.

By carefully aggregating and curating ploidy and breeding system data for Solanaceae, we showed how  the complex interaction between ploidy and breeding system  can lead to diversification conclusions that differ from the diversification conclusions of models that focus on one trait at a time.

@@

% 3) I'll bet allowing cladogenetic transitions would make the one-step ID/CP even more dominant.
%    I added the paragraph below regarding cladogenetic transitions, but it doesn't really fit with
%    the rest of the discussion, so modify or discard as needed:
% E: A small defense of excluding clado is that it didn't affect the diversification results in goldberg2010 vs goldberg2012
Our work shows that a diploid Solanaceae lineage is much more likely to take a
one-step ID/CP pathway to polyploidy rather than a two-step ID/CD/CP pathway. 
We modeled both of these pathways as anagenetic character changes occurring within a species.
However \citet{goldberg_2012} showed that I/C transitions are often associated with speciation events, and similarly \citet{freyman_2017} demonstrated that D/P transitions may also be associated with speciation events. 
Models that fail to consider transitions that occur at speciation (cladogenetic changes) may
make misleading estimates of anagenetic transition rates.
The HiSSE-based models introduced in our work here could be extended to incorporate cladogenetic transitions. 
It is possible that allowing for cladogenetic transitions may enable the one-step pathway
to even further dominate over the two-step pathway to polyploidy, but this remains to be tested in future work.

% And, are diversification rates of polyploid SC lineages different if one or other path was taken? \citet{charlesworth1985} % E: This is a great question.  We don't answer it with this round of analyses.  But discuss?
% E: To elaborate: In the IC/CD/CP model, the CP state would be subdivided into two hidded values, corresponding to the influx from the two pathways.  Those two states, CP_X and CP_Y, could have different lambda and mu.

Diploidization needs to be considered in diversification models because higher rates of net diversification (speciation minus extinction) in diploids can be obtained from models that ignore the possibility of a polyploid lineage that has diploidized being the diversification enhancer. 
Hence, in a diploidization lacking model, diploids will show higher rates of diversification when in fact it is a polyploidy event and its subsequent diploidization that generated higher speciation and less extinction.  
By adding diploidization to models of polyploidy linked to diversification it is possible to recover this complicated scenario and to reconcile genomic evidence with stochastic models. 
In the cases where ploidy level was assigned by us based chromosome multiplicity at the genus level, estimating diploidization  might be a potential issue due to ploidy misclassification.

%B: v2.0
\subsection*{Diploidization}

Diploidization is widely considered to be relatively rare, compared with polyploidization, but flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history. %cite
As a consequence, it is possible that nearly all extant species classified as ``diploid" in our analyses, based on cytogenetic data, are possibly secondary diploids, having undergone both polyploidization and re-diploidization.
Unduly ignoring secondary diploidization would clearly underestimate the rate of transitions from polyploids to diploids, and would likely cause inflated net diversification rates for diploids. % E: first clause seems like a tautology? second clause does not seem obvious, but could at least say that it could affect div rate estimates
We find strong support in models that included diploidization \cref{table:bayesfactors}. However, estimating the rate at which it occurs based on ploidy level classifications is hard and highly uncertain (see parameter $\delta$ in \cref{suppfigure:DP,suppfigure:DPAB,suppfigure:IDCDCP,suppfigure:IDCDCPAB}) and can change the pattern of ancestral reconstructions (\cref{suppfigure:DPnodipasr,suppfigure:DPasr,suppfigure:DPnodipABasr,suppfigure:DPABasr,suppfigure:IDCDCPnodipasr,suppfigure:IDCDCPasr,suppfigure:IDCDCPnodipABasr,suppfigure:IDCDCPABasr}).
%We find weak support for inclusion of diploidization. %cite fig and expand (FIXME).
% R- I corrected the above phrase, we found strong support for the inclusion of diploidization, models that have it are always preferred over models that don't however that parameter is hard to estimate

@@

On the other hand, there is little evidence for occurrence of re-diploidization in the time frame we considered, within the nightshades.
First, the Ks-inferred duplication long pre-dates the origin of the family \citep{sarkinen_2013}. 
The best-supported, most-recent WGD, which may represent a reversion from polyploidy to diploidy, is older than the lineage under consideration in this study.
At best, the species we consider are ancestrally diploid, having re-diploidized after a suspected WGD.
% E: But then how is there SI??
Second, studies comparing map-based genome synteny within the family find no evidence for recent diploidization \citep{wu_2010a}.
Instead, simple genome re-arrangements appear sufficient to explain chromosomal evolution between a number of species, including all of those in the relatively cytogenetically conserved `x=12' group, which includes tomato, potato, eggplant, pepper, and tobacco.
%B: Once this is re-arranged, and made to fit with the rest of the discussion, add what this has to do with re-diploidization and how to interpret $\delta$.
It is perhaps not surprising that models diploidization in our study recovered wide credible intervals for the diploidization parameter, $\delta$. 
Despite the limitations of assignment of ploidy levels, we found some support for models that include pararmeter $\delta$, meaning that further study may be warranted, especially in other groups.
%Moreover, rare state changes may not be easily recovered. Emma, you had a clever way to summarize this. 
%     meaning of diploidization parameter

@@

>>>>

%B:  at least two surprising outcomes: (1)  D and P on background of C, and (2) SC not as dead of an end as it was in goldberg2010, if P is partitioned out (can be framed as kind of path-dependence, i.e. qIC path different than rho path, or same as (1), so that I and C are less different on the background of D.
We investigated the individual and joint effects of ploidy and breeding system states in diversification of Solanaceae.
Most significantly, we found that the effects of ploidy, inferred as significant when it is the only trait under consideration, are overshadowed by hidden traits, or breeding system when added to models.  
Interestingly, the difference in diversification rates between the diploids and polyploids is negligible on the background of self-compatibility.
In the same vein, we recovered a smaller difference in diversification rates between self-incompatible and self-compatible lineages against the diploid genomic background.
%Inference of transition rates can appreciably across 
Motivated by the widespread findings of paleopolyploid ancestry across angiosperms, we also examined and found modest support for diploidization in the history of the family. %B: "mixed" "not strong" what's the right word here?  E: not "weak", maybe "not strong" or "not nearly as strong as anything else"
Disentangling the patterns of diversification linked to multiple traits is difficult, but our strategy allows tests of intricate hypotheses regarding a heterogenous diversification process.
Below, we discuss these key findings and highlight the most interesting and paradoxical implications.
We additionally provide an outline for broader application of phylogenetic comparative methods in the context of interactions among traits and diversification process.

\subsection*{Disentangling trait effects on diversification}

% confounding effects of other traits

% - value of approach
Disentangling the patterns of diversification linked to  multiple traits allowed us the opportunity to model and test complex hypotheses about the heterogenous diversification process in Solanaceae.
Species are created and go extinct based on multiple and often highly correlated phenotypes. 
However, estimating the speciation and extinction rates remains challenging. 
By carefully aggregating and curating ploidy and breeding system data for Solanaceae, we showed how  the complex interaction between ploidy and breeding system  can lead to diversification conclusions that differ from the diversification conclusions of models that focus on one trait at a time.

@@

\subsection*{Pathways to polyploidy}

% examine SC-only families
% no causal correlation in SSI systems
% paths to dioecy
% path-dependent net-diversification

We find that, in Solanaceae, a diploid lineage is much more likely to take a
one-step  ID$\rightarrow$CP pathway to polyploidy rather than a two-step ID$\rightarrow$CD$\rightarrow$CP pathway.
This result is significant because it has a number of implications for the inference of history of breeding and sexual systems in flowering plants.
First, support for the direct pathway lends credence to the idea that breakdown of SI by WGD---by generation of diploid `heteroalleleic' pollen---may often trigger the evolution of gender dimorphism across angiosperms \citep{miller_2000,roberston_2011}.

%missing deeper discussion of with/without SDD/SID estimates and comparison with Robertson et al. 2011.

% qualifying paragraph for pathways
% "This finding is not without a hitch..."
We implicitly modeled trait changes in both of these pathways as anagenetic, transitions occurring within lineages, without associated speciation events.
However, losses of SI (I$\rightarrow$C transitions) are often associated with cladogenesis in Solanaceae \citep{goldberg_2012}, and polyploidization (D$\rightarrow$P transition) also seem to co-occur with speciation events in a number of other families \citep{zhan_2016,freyman_2017}.
Models that fail to consider such cladogenetic changes may make misleading estimates of anagenetic transition rates.
In the only available data bearing on this problem, \citet{goldberg_2012} found that extending the *SSE framework to allow cladogenetic changes did not substantially affect inference of net diversification rates.
Nevertheless, a distinct possibility remains that allowing for cladogenetic transitions could tip the scales further in favor of the one-step pathway to SC polyploidy (CP), but this remains to be tested in future work.
Specifically, the multi/hidden state SSE-based framework we employed here could be extended to incorporate cladogenetic transitions.
% We could also list %W+%R idea above, for stochastic mapping?
Another useful line of inquiry could examine the path-dependent diversification rates, asking whether diversification rates of polyploid SC lineages differ depending on the path taken.

% And, are diversification rates of polyploid SC lineages different if one or other path was taken? \citet{charlesworth1985} % E: This is a great question.  We don't answer it with this round of analyses.  But discuss?
% E: To elaborate: In the IC/CD/CP model, the CP state would be subdivided into two hidden values, corresponding to the influx from the two pathways.  Those two states, CP_X and CP_Y, could have different lambda and mu.

%W: 3) I'll bet allowing cladogenetic transitions would make the one-step ID/CP even more dominant.
%    I added the paragraph below regarding cladogenetic transitions, but it doesn't really fit with
%    the rest of the discussion, so modify or discard as needed:
% E: A small defense of excluding clado is that it didn't affect the diversification results in goldberg2010 vs goldberg2012

\subsection*{Diploidization}

%R: By adding diploidization to models of polyploidy linked to diversification it is possible to recover this complicated scenario and to reconcile genomic evidence with stochastic models. 
%In the cases where ploidy level was assigned by us based chromosome multiplicity at the genus level, estimating diploidization  might be a potential issue due to ploidy misclassification.

Diploidization is widely considered to be relatively rare, compared with polyploidization \citep{husband_2013}, but flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}.
As a consequence, it is possible that nearly all extant species classified as ``diploid" in our analyses, based on cytogenetic data, are possibly secondary diploids, having undergone both polyploidization and re-diploidization.
Unduly ignoring secondary diploidization would clearly underestimate the rate of transitions from polyploids to diploids, and would likely cause inflated net diversification rates for diploids. % E: first clause seems like a tautology? second clause does not seem obvious, but could at least say that it could affect div rate estimates
We find strong support in models that included diploidization \cref{table:bayesfactors}. However, estimating the rate at which it occurs based on ploidy level classifications is hard and highly uncertain (see parameter $\delta$ in \cref{suppfigure:DP,suppfigure:DPAB,suppfigure:IDCDCP,suppfigure:IDCDCPAB}) and can change the pattern of ancestral reconstructions (\cref{suppfigure:DPnodipasr,suppfigure:DPasr,suppfigure:DPnodipABasr,suppfigure:DPABasr,suppfigure:IDCDCPnodipasr,suppfigure:IDCDCPasr,suppfigure:IDCDCPnodipABasr,suppfigure:IDCDCPABasr}).
%We find weak support for inclusion of diploidization. %cite fig and expand (FIXME).
% R- I corrected the above phrase, we found strong support for the inclusion of diploidization, models that have it are always preferred over models that don't however that parameter is hard to estimate

@@

%B: incorporate ancestral states in absence of I/C consideration to rag on \delta
%Natural and induced tetraploids from a variety of families, including Fabaceae, Onagraceae, Rosaceae, and Solanaceae, almost invariably lose gametophytic SI (reviewed in \citealt{ramsey_1998,stone_2002,mcclure_2009}).
On the other hand, there is little evidence for occurrence of re-diploidization in the time frame we considered, within the nightshades.
First, the Ks-inferred duplication long pre-dates the origin of the family \citep{sarkinen_2013}. 
The best-supported, most-recent WGD, which may represent a reversion from polyploidy to diploidy, is older than the lineage under consideration in this study.
At best, the species we consider are ancestrally diploid, having re-diploidized after a suspected WGD.
% E: But then how is there SI?! Agreed, I'll cover it.
Second, studies comparing map-based genome synteny within the family find no evidence for recent diploidization \citep{wu_2010a}.
Instead, simple genome re-arrangements appear sufficient to explain chromosomal evolution between a number of species, including all of those in the relatively cytogenetically conserved `x=12' group, which includes tomato, potato, eggplant, pepper, and tobacco.
%B: Once this is re-arranged, and made to fit with the rest of the discussion, add what this has to do with re-diploidization and how to interpret $\delta$.
It is perhaps not surprising that models diploidization in our study recovered wide credible intervals for the diploidization parameter, $\delta$. 
Despite the limitations of assignment of ploidy levels, we found some support for models that include pararmeter $\delta$, meaning that further study may be warranted, especially in other groups.
%Moreover, rare state changes may not be easily recovered. Emma, you had a clever way to summarize this. 
%     meaning of diploidization parameter

@@

\bigskip

\subsection*{Conclusions}
% problems: this system has a big set of advantages - carefully curated (co-)occurrences of two traits, genetic basis, etc. and everything is still contradictory and opaque.

### 2019-03-12 manuscript/introduction.tex
- backups of all the changes -- mainly to discussion

Polyploidization is a remarkably common mutation in plants \citep{husband_2013, zenilferguson_2017}.
The widespread prevalence of variation in ploidy has long been considered a salient feature of flowering plant lineages \citep{stebbins1938}. 
Increase in ploidy can alter many traits and impact a variety of evolutionary \citep{ramsey_2002} and ecological processes \citep{sessa_2019}.
%
Polyploids are thought to have lower diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
Recent studies, however, find common and numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that whole genome duplications have played an important macroevolutionary role driving innovation and diversification in plants. 
Evidence of paleo-polyploidy within the genomes of diploid extant plants also implies pervasiveness of diploidization, the return of polyploids to the diploid state, throughout the angiosperm phylogeny \citep{soltis_2015, dodsworth_2015}.
This could affect the interpretation of different net diversification rates for diploids and polyploids, especially if diploids have a relatively higher net diversification rate if they were ancestrally polyploid than if they had no polyploid ancestry.

>>>>

Polyploidization is a remarkably common mutation in plants \citep{husband_2013, zenilferguson_2017}.
The widespread prevalence of variation in ploidy has long been considered a salient feature of flowering plant lineages \citep{stebbins1938}. 
Increase in ploidy can alter many traits and impact a variety of evolutionary \citep{ramsey_2002} and ecological processes \citep{sessa_2019}.
%
Polyploids are thought to have lower diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
Recent studies, however, find common and numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that whole genome duplications (WGDs) have played an important macroevolutionary role driving innovation and diversification in plants. 
Evidence of paleo-polyploidy within the genomes of diploid extant plants also implies pervasiveness of diploidization, the return of polyploids to the diploid state, throughout the angiosperm phylogeny \citep{soltis_2015, dodsworth_2015}.
This could affect the interpretation of different net diversification rates for diploids and polyploids, especially if diploids have a relatively higher net diversification rate if they were ancestrally polyploid than if they had no polyploid ancestry.

### 2019-03-12 manuscript/results.tex
- backups of all the changes -- mainly to discussion

There are two pathways by which a SI diploid lineages will eventually---given enough time---become a SC polyploid.
In the one-step pathway, polyploidization directly disables SI.
In the two-step pathway, SI is first lost within the diploid state, followed by polyploidization.
Determining the relative contribution of these pathways based on our estimated transition rates from the ID/CD/CP model, we find that the one-step pathway is more likely on short timescales and the two-step pathway is more likely on long timescales (\cref{figure:pathways}, left panels).
When not much time has elapsed, the one-step pathway is more likely because only one event need happen.
When more time has elapsed, the two-step pathway is more likely because the rate of loss of SI within diploids, $q_{IC}$, is greater than the rate of polyploidization for SI species, $\rho_I$ (\cref{suppfigure:IDCDCPnodip}).
% These conclusions are agree with those of \citet{robertson_2011}.

When rates of lineage diversification are incorporated, however, the conclusions change.
Beginning again with a single ID lineage, it is now necessary to consider its possible proliferation in the ID state, and also in the CD state for the two-step pathway, and to track the progression of the full resulting clade to the CP state.
In this situation, even over long timescales, the two-step pathway contributes less (\cref{figure:pathways}, right panels).
The lower rate of net diversification in the CD state, relative to ID, means that relatively fewer lineages are available to complete the second step of the two-step pathway.

>>>>

There are two pathways by which a SI diploid lineages will eventually---given enough time---become a SC polyploid.
In the one-step pathway, polyploidization directly disables SI.
In the two-step pathway, SI is first lost within the diploid state, followed by polyploidization.
Determining the relative contribution of these pathways based on our estimated transition rates from the ID/CD/CP model, we find that the one-step pathway is more likely on short timescales and the two-step pathway is more likely on long timescales (\cref{figure:pathways}, left panels).
Beginning with a single SI diploid (ID) lineage, when not much time has elapsed, the one-step pathway is more likely because it only necessitates a single event to effect transition to SC polyploid state.
When more time has elapsed, the two-step pathway is more likely because the rate of loss of SI within diploids, $q_{IC}$, is greater than the rate of polyploidization for SI species, $\rho_I$ (\cref{suppfigure:IDCDCPnodip}).
These conclusions, however, ignore the changes in numbers of lineages in a particular state from speciation and extinction events.
By analogy, frequent sinking of a middle stepping stone in a path would lower the value of such a path.
We therefore examined the effects of diversification rate on relative importance of each pathway.
%B: revisit this possibly terrible analogy.

Indeed, incorporation of lineage diversification rates changes conclusions regarding the relative importance of each pathway.
Beginning again with a single ID lineage, it is now necessary to consider its possible proliferation in the ID state, and also in the CD state for the two-step pathway, as well as to track the progression of the full resulting clade to the CP state.
In this situation, even over long timescales, the two-step pathway contributes less (\cref{figure:pathways}, right panels).
The lower rate of net diversification in the CD state, relative to ID, means that relatively fewer lineages are available to complete the second step of the two-step pathway, ending in CP.
